---
title: "DS311 - R Lab Assignment"
author: "Your Name"
date: "8/22/2022"
output: 
  html_document:
    theme: united
    highlight: tango
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Assignment 1

* In this assignment, we are going to apply some of the build in data set in R for descriptive statistics analysis.
* To earn full grade in this assignment, students need to complete the coding tasks for each question to get the result.
* After finished all the questions, knit the document into HTML format for submission.

### Question 1

Using the **mtcars** data set in R, please answer the following questions.

```{r}
# Loading the data
data(mtcars)
install.packages('plyr', repos = "http://cran.us.r-project.org")

install.packages("lifecycle", repos = "http://cran.us.r-project.org")

library("lifecycle")

library(rlang)
library(dplyr)

install.packages("dplyr", repos = "http://cran.us.r-project.org")
# Head of the data set
head(mtcars)
```

a. Report the number of variables and observations in the data set.

```{r}
# Enter your code here!
dim(mtcars)

# Answer:
print("There are total of __11___ variables and __32___ observations in this data set.")
```



b. Print the summary statistics of the data set and report how many discrete and continuous variables are in the data set.

```{r}
# Enter your code here!
summary(mtcars)

# Answer:
print("There are _2____ discrete variables and __8___ continuous variables in this data set.")
```



c. Calculate the mean, variance, and standard deviation for the variable **mpg** and assign them into variable names m, v, and s. Report the results in the print statement.

```{r}
# Enter your code here!
mean <- mean(mtcars$mpg)
vari <- var(mtcars$mpg)
sdev <- (mtcars$mpg)



# print(paste("The average of Mile Per Gallon from this data set is ",mean , " with variance ", vari , " and standard deviation", sdev , "."))
```



d. Create two tables to summarize 1) average mpg for each cylinder class and 2) the standard deviation of mpg for each gear class.

```{r}
# Enter your code here!

install.packages("qwraps2", repos = "http://cran.us.r-project.org")

library("qwraps2")

library("tidyverse")

mtcars %>% group_by(cyl) %>% summarize((Mean = mean(mpg)))

mtcars %>% group_by(gear) %>% summarize((SDD = sd(mpg)))


```



e. Create a crosstab that shows the number of observations belong to each cylinder and gear class combinations. The table should show how many observations given the car has 4 cylinders with 3 gears, 4 cylinders with 4 gears, etc. Report which combination is recorded in this data set and how many observations for this type of car.

```{r}
# Enter your code here!
library(tidyverse)
mtcars %>%
select(cyl, gear) %>%
table()

print("The most common car type in this data set is car with __3__ cylinders and __8__ gears. There are total of __12___ cars belong to this specification in the data set.")
```


*** 

### Question 2

Use different visualization tools to summarize the data sets in this question.

a. Using the **PlantGrowth** data set, visualize and compare the weight of the plant in the three separated group. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your findings.

```{r}
# Load the data set
data("PlantGrowth")

# Head of the data set
head(PlantGrowth)

# Enter your code here!
install.packages("tidyverse", repos =  "http://cran.us.r-project.org")

library(tidyverse)
PlantGrowth %>%
  ggplot(aes(x = group, y = weight, fill = group)) +geom_bar(stat = "identity") + labs(title = "weight of Plants in Separate groups", 
       x = "group",
      y = "weight")
```

Result:

=> Report a paragraph to summarize your findings from the plot!



b. Using the **mtcars** data set, plot the histogram for the column **mpg** with 10 breaks. Give labels to the title, x-axis, and y-axis on the graph. Report the most observed mpg class from the data set.

```{r}
colors <- c("Blue", "red", "orange", "grey", "purple")

hist(mtcars$mpg,
     col=colors,
     main="vehicles",
     breaks=10,
     xlin = range(10:35),
     xlab= "MPG",
     ylab="# of vehicles")


print("Most of the cars in this data set are in the class of ___15___ mile per gallon.")
```



c. Using the **USArrests** data set, create a pairs plot to display the correlations between the variables in the data set. Plot the scatter plot with **Murder** and **Assault**. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your results from both plots.

```{r}
# Load the data set
data("USArrests")

# Head of the data set
head(USArrests)

# Enter your code here!
plot(y = USArrests$Murder, x = USArrests$Assault, main = "Murder Rate vs. Assault Rate",xlab = "Assult")

```

Result:

=> Report a paragraph to summarize your findings from the plot!



***

### Question 3

Download the housing data set from www.jaredlander.com and find out what explains the housing prices in New York City.

Note: Check your working directory to make sure that you can download the data into the data folder. 

```{r, echo=FALSE}
# Load and clean the housing data set 
download.file(url='https://www.jaredlander.com/data/housing.csv',
              destfile='data/housing.csv', mode='wb')
housingData <- read.csv('data/housing.csv')
housingData <- subset(housingData, 
                      select = c("Neighborhood", "Market.Value.per.SqFt", "Boro", "Year.Built"))
housingData <- na.omit(housingData)                   
```

a. Create your own descriptive statistics and aggregation tables to summarize the data set and find any meaningful results between different variables in the data set.

```{r}
# Head of the cleaned data set
head(housingData)

# Enter your code here!
summary(housingData)
```

b. Create multiple plots to demonstrates the correlations between different variables. Remember to label all axes and give title to each graph.

```{r}
# Enter your code here!
library(RColorBrewer)
library(ggplot2)
library(tidyverse)
 plot(y = housingData$Year.Built, x = housingData$Market.Value.per.SqFt, main = "Market. Value vs. year.B")
```

c. Write a summary about your findings from this exercise.

=> Enter your answer here!
I found that cheaper houses are the older ones and newer houses are the ones that sell faster 

